%
%  $Description: Author guidelines and sample document in LaTeX 2e$ 
%

\documentclass[10pt,twocolumn]{article} 
\usepackage{times,mathptmx,fullpage}
% Allows you to put literal text into your document without having
% LaTeX try to interpret it.
\usepackage{verbatim}
% Sorts citations numerically and "combines" adjacent ones
\usepackage{cite}
% Allows you to individually label subfigures in a multi-part figure
\usepackage{subfigure}
% Allows the definition of macros that are smart about adding space
% after text in a macro
\usepackage{xspace}

\usepackage{graphicx}

% Redefine the percentage of the page that can be used for floats (figures,
% tables, etc.)
\renewcommand\floatpagefraction{.9}
\renewcommand\dblfloatpagefraction{.9}
\renewcommand\topfraction{.9}
\renewcommand\dbltopfraction{.9}
\renewcommand\bottomfraction{.9}
\renewcommand\textfraction{.1}
\setcounter{totalnumber}{10}
\setcounter{topnumber}{10}
\setcounter{dbltopnumber}{10}
\setcounter{bottomnumber}{10}

% Set values for float separation from text
\setlength{\floatsep}{1.5ex plus1.0ex minus 0.2ex}
\setlength{\dblfloatsep}{1.5ex plus1.0ex minus 0.2ex}
\setlength{\textfloatsep}{1.5ex plus1.0ex minus 0.2ex}
\setlength{\dbltextfloatsep}{1.5ex plus1.0ex minus 0.2ex}
\setlength{\abovecaptionskip}{0.5ex}
\setlength{\belowcaptionskip}{0.5ex}

% Don't allow widows or clubs - single lines at the start/end of a column
\widowpenalty=10000
\clubpenalty=10000

\newcommand{\latex}{\LaTeX\xspace}
\pagestyle{plain}

%------------------------------------------------------------------------- 
\begin{document}

\title{Developing Disposable Wireless Sensor Networks}

\author{
Zak Rubin \\
\textit{Univ. of California, Santa Cruz} \\
\textit{zrubin@soe.ucsc.edu}
\and
H. Blake Skinner \\
\textit{Univ. of California, Santa Cruz} \\
\textit{hskinner@soe.ucsc.edu}
}

\maketitle
\thispagestyle{empty}

\begin{abstract}
Ad-hoc wireless sensor networks are predicted to increase in use significantly as advances in their core
hardware components continue.  They have already attracted considerable interest in academia and have
been successfully deployed for a variety of applications.  Despite advances in this area, cost and power
limitations remain the major obstacles to their use.

In this paper we explore the general problem of transferring data around a network
at minimal cost.  We present several protocols to transfer data across an ad-hoc network while
preserving the data's locality.  We will explore the problem of exchanging throughput for power
savings and will demonstrate what is possible in terms of through-put and power use with hardware
currently available on the market.  With this we hope to demonstrate the cost trade-offs associated
with wireless ad-hoc networks today how throughput can be sacrificed for power savings.
\end{abstract}

\input{introduction}
\input{related_work}
\input{design_considerations}
\input{gumbo_node}
\input{evaluation}
\input{challenges}
\input{future_work}
\input{acknowledgements}

\bibliographystyle{latex8}
\bibliography{gumbo_bib}

\end{document}
